Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-11-30 paper/paper.tex
Some of Ruth's things

% cref style
\crefname{figure}{Figure}{Figures}
\Crefname{figure}{Figure}{Figures}
\crefname{section}{Section}{Sections}
\Crefname{section}{Section}{Sections}
\crefname{proposition}{Proposition}{Propositions}
\Crefname{proposition}{Proposition}{Propositions}

@@

\begin{proposition}
    \label{prop:comp}
Let $i$ be an assignment of $T$ vertices to $P$ vertices.  Then $i$ satisfies the definition of $P
\hookrightarrow T$ if and only if $i$ satisfies $P \rightarrowtail T$ and
$\loopcomp{P} \rightarrowtail \loopcomp{T}$.
Similarly, $i$ satisfies the definition of $P
\lessind{k} T$ if and only if $i$ satisfies both $P \lessnonind{k} T$ and
$\loopcomp{P}
\lessnonind{k} \loopcomp{T}$.
\end{proposition}

@@

For both subgraph isomorphism and maximum common subgraph, constraint programming is the best known
approach\footnote{Note that toolkits are not used in state of the art implementations, to facilitate
better memory layouts and propagation queues, although the algorithms are inspired by
constraint programming.}, although a reduction to maximum clique is better when edge labels are
present \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}. For both problems, we create a variable for each vertex
in the pattern graph (the smaller graph, in the case of maximum common subgraph), with domains
ranging over the vertices in the target graph (the larger graph). For maximum common subgraph, each
domain is given an additional $\bot$ value, meaning ``unmapped''.

@@

There are three easy ways we might try to extend existing algorithms for this problem.
A non-induced $k$-less subgraph isomorphism from $P$ to $T$ is equivalent to a subgraph isomorphism
between $P$ and $T$, with $k$ extra universally-adjacent vertices added to $T$, and so we could try
solving subgraph isomorphism with a modified target graph. However, using such
an approach is not ideal, because it would introduce symmetries; for induced $k$-less subgraph
isomorphisms, an approach based around adding vertices to $T$ cannot work at all.
Another algorithmic approach could be to try each way of removing $k$ vertices from the pattern
graph, and solving each subgraph isomorphism problem in turn. This approach might be feasible for $k
= 1$, although it would involve a lot of duplication of search effort, but for larger values of $k$
the number of searches which would have to be made would grow very quickly.
Finally, for the induced case we could try adapting maximum common subgraph algorithms to solve the
decision problem. However, both the maximum common subgraph algorithm and the clique encoding
require $O(\left|\V(P)\right|^2\left|\V(T)\right|^2)$ memory \citep{DBLP:conf/cp/McCreeshNPS16},
which is extremely problematic on the instances with which we wish to work.

@@

        \path (p) edge (a) edge (c)
            (a) edge (d)
            (c) edge (d);
        \end{scope}
    \end{tikzpicture}
    \caption{There is a non-induced isomorphism $P\rightarrowtail T$ from the first graph to the
    second, using the highlighted vertices. No such isomorphism exists to the third graph, but there
    is a $1$-less isomorphism $P\lessnonind{1} U$ omitting vertex $b$.}\label{figure:klessexample}
\end{figure}

@@

\Cref{figure:klessexample} also illustrates the three cases possible when filtering by neighbourhood
degree sequence.  Removing vertex $d$ causes each entry in $\nds(p)$ to be reduced; removing vertex
$a$ removes an entry from $\nds(p)$; and removing either of vertex $b$ or $c$ in the middle layer
causes both the size of $\nds(p)$ to be reduced and an entry in $\nds(p)$ to be reduced.

@@

\Cref{figure:ids} demonstrates that this is effective in practice: we show the amount of domain
reduction we can achieve at the top of search by using invariants and a fixed $k$, compared to the
search space size for the maximum common induced subgraph problem. (The results on the non-induced
version show a similar but slightly weaker trend.) We look at the product of the domain sizes,
rather than the number of eliminations, as this better reflects the number of combinations remaining
to be considered.

@@

We begin by trying to infer domain deletions. The $\FuncSty{propagate}$ function looks for domains
which contains either only a single value, or only wildcards---we call such a domain
``effectively-unit''. If such a domain $D_v$ exists, we eliminate its value from every other domain,
and then propagate adjacency: for each domain corresponding to a vertex adjacent to $v$, we
eliminate any value from its domain which is not adjacent to $v$ (treating wildcards as being
adjacent to all vertices). If a domain wipeout occurs, we return failure.

@@

If no effectively-unit domains remain, we attempt stronger propagation for the all-different
constraint. Our $\FuncSty{allDifferent}$ function is the bit-parallel propagator taken from
\citet{DBLP:conf/cp/McCreeshP15}, and is \emph{not} the usual matching-based propagator
\citep{DBLP:conf/aaai/Regin94} which guarantees generalised arc consistency. Because we use multiple
wildcard values, we do not need to modify the algorithm to allow a single wildcard value to be used
more than once. This propagation could create new effectively-unit domains; if so, we loop.

>>>>

% cref style
\crefname{figure}{Figure}{Figures}
\Crefname{figure}{Figure}{Figures}
\crefname{section}{Section}{Sections}
\Crefname{section}{Section}{Sections}
\crefname{proposition}{Proposition}{Propositions}
\Crefname{proposition}{Proposition}{Propositions}
\crefname{corollary}{Corollary}{Corollaries}
\Crefname{corollary}{Corollary}{Corollaries}

@@

\begin{proposition}
    \label{prop:comp}
Let $i$ be an assignment of vertices of $T$ to vertices of $P$.  Then $i$ satisfies the definition of $P
\hookrightarrow T$ if and only if $i$ satisfies $P \rightarrowtail T$ and
$\loopcomp{P} \rightarrowtail \loopcomp{T}$.
Similarly, $i$ satisfies the definition of $P
\lessind{k} T$ if and only if $i$ satisfies both $P \lessnonind{k} T$ and
$\loopcomp{P}
\lessnonind{k} \loopcomp{T}$.
\end{proposition}

@@

For both subgraph isomorphism and maximum common subgraph, constraint programming is the best known
approach\footnote{Note that toolkits are not used in state of the art implementations, to facilitate
better memory layouts and propagation queues, although the algorithms are inspired by constraint
programming.}, although a reduction to the maximum clique problem is better when edge labels are
present \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}. For both problems, we create a
variable for each vertex in the pattern graph (the smaller graph, in the case of maximum common
subgraph), with domains ranging over the vertices in the target graph (the larger graph). For
maximum common subgraph, each domain is given an additional $\bot$ value, meaning ``unmapped''.

@@

There are three easy ways we might try to extend existing algorithms to handle the $k$-less problem.
A non-induced $k$-less subgraph isomorphism from $P$ to $T$ is equivalent to a subgraph isomorphism
between $P$ and $T$, with $k$ extra universally-adjacent vertices added to $T$, and so we could try
solving subgraph isomorphism with a modified target graph. However, using such
an approach is not ideal, because it would introduce symmetries; for induced $k$-less subgraph
isomorphisms, an approach based around adding vertices to $T$ cannot work at all.
Another algorithmic approach could be to try each way of removing $k$ vertices from the pattern
graph, and solving each subgraph isomorphism problem in turn. This approach might be feasible for $k
= 1$, although it would involve a lot of duplication of search effort, but for larger values of $k$
the number of searches which would have to be made would grow very quickly.
Finally, for the induced case we could try adapting maximum common subgraph algorithms to solve the
decision problem. However, both the maximum common subgraph algorithm and the clique encoding
require $O(\left|\V(P)\right|^2\left|\V(T)\right|^2)$ memory \citep{DBLP:conf/cp/McCreeshNPS16},
which is extremely problematic on the instances with which we wish to work.

@@

        \path (p) edge (a) edge (c)
            (a) edge (d)
            (c) edge (d);
        \end{scope}
    \end{tikzpicture}
    \caption{There is a non-induced isomorphism $P\rightarrowtail T$ from the first graph to the
    second, using the highlighted vertices. No such isomorphism exists between $P$ and $U$, but there
    is a $1$-less isomorphism $P\lessnonind{1} U$ which omits vertex $b$.}\label{figure:klessexample}
\end{figure}

@@

\Cref{figure:klessexample} also illustrates the three cases possible when filtering by neighbourhood
degree sequence.  Removing vertex $d$ causes each entry in $\nds(p)$ to be reduced; removing vertex
$a$ removes an entry from $\nds(p)$; and removing either of vertex $b$ or $c$ causes both the size
of $\nds(p)$ to be reduced and an entry in $\nds(p)$ to be reduced.

@@

\Cref{figure:ids} demonstrates that \cref{prop:nds} is effective in practice: we show the amount of
domain reduction we can achieve at the top of search by using invariants and a fixed $k$, compared
to the search space size for the maximum common induced subgraph problem. (The results on the
non-induced version show a similar but slightly weaker trend.) We look at the product of the domain
sizes, rather than the number of eliminations, as this better reflects the number of combinations
remaining to be considered.

@@

We begin by trying to infer domain deletions. The $\FuncSty{propagate}$ function looks for domains
which contains either only a single value, or only wildcards---we call such a domain
\emph{effectively-unit}. If such a domain $D_v$ exists, we eliminate its value from every other domain,
and then propagate adjacency: for each domain corresponding to a vertex adjacent to $v$, we
eliminate any value from its domain which is not adjacent to $v$ (treating wildcards as being
adjacent to all vertices). If a domain wipeout occurs, we return failure.

@@

If no effectively-unit domains remain, we attempt stronger propagation for the all-different
constraint. Our $\FuncSty{allDifferent}$ function is the bit-parallel propagator taken from
\citet{DBLP:conf/cp/McCreeshP15}, and is \emph{not} the usual matching-based propagator
\citep{DBLP:conf/aaai/Regin94} which guarantees generalised arc consistency. Because we use multiple
wildcard values, we do not need to modify the algorithm to allow a single wildcard value to be used
more than once. This propagation could create new effectively-unit domains; if so, we repeat the
process.
